---
title: "Templeton Empathy Analysis"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up & Descriptive Statistics
### Set up workspace and run correlation

``` {r}

# install.packages("tidyverse", "corrplot", "ggcorrplot", "lme4", "mediation", "here", "reshape2")
require(tidyverse)
require(sjstats)
require(sjPlot)
require(lmerTest)
require(mediation)
require(here)
require(reshape2)

# Clear workspace
rm(list=ls())

empathy_data <- read.csv(here::here("templeton_empathy_clean_data.csv"))

# Helper function to calculate the mean and the standard deviation for each group ####
# Created by http://www.sthda.com/english/wiki/ggplot2-line-plot-quick-start-guide-r-software-and-data-visualization

#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
#to be summariezed
# groupnames : vector of column names to be used as
# grouping variables

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- plyr::rename(data_sum, c("mean" = varname))
  return(data_sum)
}


# Find number of children by age and country
empathy_data %>%
  group_by(age_years, country, gender) %>%
  dplyr::summarise(n = n())

# Check if empathy behaviors correlate for each country
cor.test(filter(empathy_data, country=="canada" & age_years == 6)$empathy_pain_resp_c, filter(empathy_data, country=="canada" & age_years == 6)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="canada" & age_years == 7)$empathy_pain_resp_c, filter(empathy_data, country=="canada" & age_years == 7)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="canada" & age_years == 8)$empathy_pain_resp_c, filter(empathy_data, country=="canada" & age_years == 8)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="canada" & age_years == 9)$empathy_pain_resp_c, filter(empathy_data, country=="canada" & age_years == 9)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="canada" & age_years == 10)$empathy_pain_resp_c, filter(empathy_data, country=="canada" & age_years == 10)$empathy_sorry_resp_c, use="complete.obs")

cor.test(filter(empathy_data, country=="china" & age_years == 6)$empathy_pain_resp_c, filter(empathy_data, country=="china" & age_years == 6)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="china" & age_years == 7)$empathy_pain_resp_c, filter(empathy_data, country=="china" & age_years == 7)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="china" & age_years == 8)$empathy_pain_resp_c, filter(empathy_data, country=="china" & age_years == 8)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="china" & age_years == 9)$empathy_pain_resp_c, filter(empathy_data, country=="china" & age_years == 9)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="china" & age_years == 10)$empathy_pain_resp_c, filter(empathy_data, country=="china" & age_years == 10)$empathy_sorry_resp_c, use="complete.obs")

cor.test(filter(empathy_data, country=="jordan")$empathy_pain_resp_c, filter(empathy_data, country=="jordan")$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="jordan" & age_years == 6)$empathy_pain_resp_c, filter(empathy_data, country=="jordan" & age_years == 6)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="jordan" & age_years == 7)$empathy_pain_resp_c, filter(empathy_data, country=="jordan" & age_years == 7)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="jordan" & age_years == 8)$empathy_pain_resp_c, filter(empathy_data, country=="jordan" & age_years == 8)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="jordan" & age_years == 9)$empathy_pain_resp_c, filter(empathy_data, country=="jordan" & age_years == 9)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="jordan" & age_years == 10)$empathy_pain_resp_c, filter(empathy_data, country=="jordan" & age_years == 10)$empathy_sorry_resp_c, use="complete.obs")

cor.test(filter(empathy_data, country=="south_africa")$empathy_pain_resp_c, filter(empathy_data, country=="south_africa")$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="south_africa" & age_years == 6)$empathy_pain_resp_c, filter(empathy_data, country=="south_africa" & age_years == 6)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="south_africa" & age_years == 7)$empathy_pain_resp_c, filter(empathy_data, country=="south_africa" & age_years == 7)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="south_africa" & age_years == 8)$empathy_pain_resp_c, filter(empathy_data, country=="south_africa" & age_years == 8)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="south_africa" & age_years == 9)$empathy_pain_resp_c, filter(empathy_data, country=="south_africa" & age_years == 9)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="south_africa" & age_years == 10)$empathy_pain_resp_c, filter(empathy_data, country=="south_africa" & age_years == 10)$empathy_sorry_resp_c, use="complete.obs")

cor.test(filter(empathy_data, country=="turkey")$empathy_pain_resp_c, filter(empathy_data, country=="turkey")$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="turkey" & age_years == 6)$empathy_pain_resp_c, filter(empathy_data, country=="turkey" & age_years == 6)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="turkey" & age_years == 7)$empathy_pain_resp_c, filter(empathy_data, country=="turkey" & age_years == 7)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="turkey" & age_years == 8)$empathy_pain_resp_c, filter(empathy_data, country=="turkey" & age_years == 8)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="turkey" & age_years == 9)$empathy_pain_resp_c, filter(empathy_data, country=="turkey" & age_years == 9)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="turkey" & age_years == 10)$empathy_pain_resp_c, filter(empathy_data, country=="turkey" & age_years == 10)$empathy_sorry_resp_c, use="complete.obs")

cor.test(filter(empathy_data, country=="usa")$empathy_pain_resp_c, filter(empathy_data, country=="usa")$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="usa" & age_years == 6)$empathy_pain_resp_c, filter(empathy_data, country=="usa" & age_years == 6)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="usa" & age_years == 7)$empathy_pain_resp_c, filter(empathy_data, country=="usa" & age_years == 7)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="usa" & age_years == 8)$empathy_pain_resp_c, filter(empathy_data, country=="usa" & age_years == 8)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="usa" & age_years == 9)$empathy_pain_resp_c, filter(empathy_data, country=="usa" & age_years == 9)$empathy_sorry_resp_c, use="complete.obs")
cor.test(filter(empathy_data, country=="usa" & age_years == 10)$empathy_pain_resp_c, filter(empathy_data, country=="usa" & age_years == 10)$empathy_sorry_resp_c, use="complete.obs")

#Visualize correlations between behaviors

empathy_data %>%
  ggplot(aes(x=empathy_pain_resp, y=empathy_sorry_resp, group=country, color=country)) + 
  geom_smooth() +
  geom_point()+
  scale_color_brewer(palette="Paired")+theme_minimal() +
  xlab("Cognitive Empathy Score") + ylab("Empathic Concern Score") +
  labs(colour = "Country") +
  facet_wrap(~country) +
  theme(plot.title = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 20, angle = 90, hjust = .5, vjust = .5, face = "plain"))

```

```{r}
head(empathy_data)
```

# Empathy Behavior and Cultural Collectivism
#### Hypothesis: If a child is raised in a more collectivist country, then they will show greater cognitive and affective empathy, replicating the findings of Chopik, O'Brian & Konrath 2016
Dependent Variables: Child behaviors (cognitive empathy and empathic concern) 
Independent Variables: High-low median split based on Hofstede individualism score
Analyses: For each behavior, t-test between collectivist and individualist countries
Interpretation: If behavior scores are higher in high-collectivism countries, then our data replicate the Chopik findings that those in more collectivist countries show greater empathy


```{r}
t.test(filter(empathy_data, individualism_split=="individualist")$empathy_pain_resp, filter(empathy_data, individualism_split=="collectivist")$empathy_pain_resp)

t.test(filter(empathy_data, individualism_split=="individualist")$empathy_pain_resp_c, filter(empathy_data, individualism_split=="collectivist")$empathy_pain_resp_c)

t.test(filter(empathy_data, individualism_split=="individualist")$empathy_sorry_resp, filter(empathy_data, individualism_split=="collectivist")$empathy_sorry_resp)

t.test(filter(empathy_data, individualism_split=="individualist")$empathy_sorry_resp_c, filter(empathy_data, individualism_split=="collectivist")$empathy_sorry_resp_c)

# Visualize parent QCAE by split
parent_summary <- reshape2::melt(empathy_data %>%
  group_by(individualism_split) %>%
  dplyr::summarise(parent_aff = mean(qcaeparent_aff, na.rm=TRUE), parent_cog = mean(qcaeparent_cog, na.rm=TRUE), parent_tot = mean(qcaeparent_tot, na.rm=TRUE )), id.vars="individualism_split")

ggplot(parent_summary, aes(fill=individualism_split, y=value, x=variable)) + 
    geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("parent_cog" = "QCAE Cognitive Empathy", "parent_aff" = "QCAE Affective Empathy", "parent_tot" = "QCAE Total Empathy")) + 
  ylab("Parent Score") +
  theme(axis.title.x = element_blank()) +
  labs(fill = "Cultural Split")
  
# Visualize child QCAE by split
child_summary <- melt(empathy_data %>%
  group_by(individualism_split) %>%
  dplyr::summarise(child_aff = mean(qcaechild_aff, na.rm=TRUE), child_cog = mean(qcaechild_cog, na.rm=TRUE), child_tot = mean(qcaechild_tot, na.rm=TRUE)), id.vars="individualism_split")

ggplot(child_summary, aes(fill=individualism_split, y=as.numeric(value), x=variable)) + 
    geom_bar(position="dodge", stat="identity") +
  ylim(NA, 100) +
  scale_x_discrete(labels=c("child_cog" = "QCAE Cognitive Empathy", "child_aff" = "QCAE Affective Empathy",
                              "child_tot" = "QCAE Total Empathy")) + 
  ylab("Child Score") +
  theme(axis.title.x = element_blank()) +
  labs(fill = "Cultural Split")
  
# Visualize child QCAE by split
behavior_summary <- melt(empathy_data %>%
  group_by(individualism_split) %>%
  dplyr::summarise(cog_emp = mean(empathy_pain_resp, na.rm=TRUE), emp_con = mean(empathy_sorry_resp, na.rm=TRUE)), id.vars="individualism_split")

ggplot(behavior_summary, aes(fill=individualism_split, y=as.numeric(value), x=variable)) + 
    geom_bar(position="dodge", stat="identity") +
  ylim(NA, 100) +
  scale_x_discrete(labels=c("cog_emp" = "Cognitive Empathy Score", "emp_con" = "Empathic Concern")) + 
  ylab("Child Score") +
  theme(axis.title.x = element_blank()) +
  labs(fill = "Cultural Split")


```


# Self-Reported Empathy and Cultural Collectivism
#### Hypothesis: If a child is raised in a more collectivist country, then they will show greater cognitive and affective empathy, replicating the findings of Chopik, O'Brian & Konrath 2016
Dependent Variables: QCAE scores (affective, cognitive, and total for child and parent)
Independent Variables: High-low median split based on Hofstede individualism score
Analyses: For each child QCAE score, t-test between collectivist and individualist countries
Interpretation: If QCAE scores are higher in high-collectivism countries, then our data replicate the Chopik findings that those in more collectivist countries show greater empathy


```{r}
t.test(filter(empathy_data, individualism_split=="individualist")$qcaeparent_cog, filter(empathy_data, individualism_split=="collectivist")$qcaeparent_cog)

t.test(filter(empathy_data, individualism_split=="individualist")$qcaeparent_aff, filter(empathy_data, individualism_split=="collectivist")$qcaeparent_aff)

t.test(filter(empathy_data, individualism_split=="individualist")$qcaeparent_tot, filter(empathy_data, individualism_split=="collectivist")$qcaeparent_tot)

t.test(filter(empathy_data, individualism_split=="individualist")$qcaechild_cog, filter(empathy_data, individualism_split=="collectivist")$qcaechild_cog)

t.test(filter(empathy_data, individualism_split=="individualist")$qcaechild_aff, filter(empathy_data, individualism_split=="collectivist")$qcaechild_aff)

t.test(filter(empathy_data, individualism_split=="individualist")$qcaechild_tot, filter(empathy_data, individualism_split=="collectivist")$qcaechild_tot)
```

# Development of Empathy
####	Hypothesis: Cognitive empathy and empathic concern will vary by child age
Dependent Variables: Age in months, age in years
Independent Variables: Child behaviors (cognitive empathy and empathic concern)
Analyses: Linear regressions predicting each behavior by each age type
Interpretation: If there is a significant main effect of age on each behavior, then empathic concern and cognitive empathy vary by age


``` {r, echo = TRUE}
pain_years <- lm(empathy_pain_resp_c ~ age_years, data = empathy_data)
summary(pain_years)
pain_months <- lm(empathy_pain_resp_c ~ age_months, data = empathy_data)
summary(pain_months)

sorry_years <- lm(empathy_sorry_resp_c ~ age_years, data = empathy_data)
summary(sorry_years)
sorry_months <- lm(empathy_sorry_resp_c ~ age_months, data = empathy_data)
summary(sorry_months)

ggplot(data=empathy_data, aes(x=age_months, y=empathy_pain_resp)) +
  geom_point()+
  geom_smooth(method = lm) + 
  labs(x = "Age (months)", y = "Cognitive Empathy")
  

 ggplot(empathy_data, aes(x=as.factor(age_years), y=empathy_pain_resp)) + 
   geom_boxplot( )

```


#### Empathy’s Developmental Trajectory by Country
#	Hypothesis: Empathy score trajectories will vary by country/ gender
Dependent Variables: Empathy scores (cognitive empathy and empathic concern scores)
Independent Variables: age_months, age_years, country, gender
Analyses: For each type of age variable, run the following:
1.	lm(behavior ~ age * country)
2.	lm(behavior ~ age * country + gender)
3.	Compare these two using AIC scores
Interpretation: If the interaction of age and country is significant, then trajectories vary by country. If model 2 has a lower AIC score, then the behavior also varies by gender.


``` {r, echo = TRUE}
pain_years_country <- lm(empathy_pain_resp_c ~ age_years * country, data = empathy_data)
summary(pain_years_country)
pain_years_country_gender <- lm(empathy_pain_resp_c ~ age_years * country + gender, data = empathy_data)
summary(pain_years_country_gender)
AIC(pain_years_country)
AIC(pain_years_country_gender) # adding gender improves fit

pain_months_country <- lm(empathy_pain_resp_c ~ age_months * country, data = empathy_data)
summary(pain_months_country)
pain_months_country_gender <- lm(empathy_pain_resp_c ~ age_months * country + gender, data = empathy_data)
summary(pain_months_country_gender)
AIC(pain_months_country)
AIC(pain_months_country_gender) # adding gender improves fit

sorry_years_country <- lm(empathy_sorry_resp_c ~ age_years * country, data = empathy_data)
summary(sorry_years_country)
sorry_years_country_gender <- lm(empathy_sorry_resp_c ~ age_years * country + gender, data = empathy_data)
summary(sorry_years_country_gender)
AIC(sorry_years_country)
AIC(sorry_years_country_gender) # adding gender improves fit

sorry_months_country <- lm(empathy_sorry_resp_c ~ age_months * country, data = empathy_data)
summary(sorry_months_country)
sorry_months_country_gender <- lm(empathy_sorry_resp_c ~ age_months * country + gender, data = empathy_data)
summary(sorry_months_country_gender)
AIC(sorry_months_country)
AIC(sorry_months_country_gender) # adding gender improves fit

# Summarize and visualize best-fitting models

plot_model(pain_years_country_gender, type='int') 
tab_model(pain_years_country_gender, show.std = 'std')
plot_model(pain_months_country_gender, type='int') 
tab_model(pain_months_country_gender, show.std = 'std')

plot_model(sorry_years_country_gender, type='int') 
tab_model(sorry_years_country_gender, show.std = 'std')
plot_model(sorry_months_country_gender, type='int') 
tab_model(sorry_months_country_gender, show.std = 'std')

# Visualization of each behavior by age, country, and gender

pain_years_summary <- data_summary(empathy_data, varname="empathy_pain_resp", groupnames=c("country", "age_months", "gender"))

pain_years_summary %>%
  filter(!is.na(gender)) %>%
  ggplot(aes(x=age_months, y=empathy_pain_resp, group=country, color=country)) + 
  geom_point()+
  geom_smooth(method = lm) +
  scale_color_brewer(palette="Paired")+theme_minimal() +
  xlab("Age (years)") + ylab("Cognitive Empathy Score") +
  labs(colour = "Country") +
  facet_wrap(~gender) +
  theme(plot.title = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 20, angle = 90, hjust = .5, vjust = .5, face = "plain"))

sorry_years_summary <- data_summary(empathy_data, varname="empathy_sorry_resp", groupnames=c("country", "age_years", "gender"))

sorry_years_summary %>%
  filter(!is.na(gender)) %>%
  ggplot(aes(x=age_years, y=empathy_sorry_resp, group=country, color=country)) + 
  geom_line() + geom_point()+
  scale_color_brewer(palette="Paired")+theme_minimal() +
  xlab("Age (years)") + ylab("Cognitive Empathy Score") +
  labs(colour = "Country") +
  facet_wrap(~gender) +
  theme(plot.title = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 20, angle = 90, hjust = .5, vjust = .5, face = "plain"))

```

#### Parent and Child Empathy
# Hypothesis: Parent-reported cognitive and affective empathy will better predict empathic concern and cognitive empathy behaviors than age and country alone
Dependent Variables: Child behaviors (empathic concern and cognitive empathy)
Independent Variables: Total QCAE score for parent, age in months, age in years, country
Analyses: For each type of age variable, run the following:
1. lm(behavior ~ age * country + QCAEparent_tot)
2.	Compare this to lm(behavior ~ age) using AIC scores
Interpretation: If model 2 has a lower AIC score, then empathy score varies by parent QCAE.


``` {r, echo = TRUE}
pain_years_country_qcae <- lm(empathy_pain_resp_c ~ age_years * country + qcaeparent_tot, data = empathy_data)
summary(pain_years_country_qcae)
AIC(pain_years_country) 
AIC(pain_years_country_qcae) # adding qcae improves fit

pain_months_country_qcae <- lm(empathy_pain_resp_c ~ age_months * country + qcaeparent_tot, data = empathy_data)
summary(pain_months_country_qcae)
AIC(pain_months_country)
AIC(pain_months_country_qcae) # adding qcae improves fit

sorry_years_country_qcae <- lm(empathy_sorry_resp_c ~ age_years * country + qcaeparent_tot, data = empathy_data)
summary(sorry_years_country_qcae)
AIC(sorry_years_country)
AIC(sorry_years_country_qcae) # adding qcae improves fit

sorry_months_country_qcae <- lm(empathy_sorry_resp_c ~ age_months * country + qcaeparent_tot, data = empathy_data)
summary(sorry_months_country_qcae)
AIC(sorry_months_country)
AIC(sorry_months_country_qcae) # adding qcae improves fit


pain_years_country_qcae_summary <- data_summary(empathy_data, varname="empathy_pain_resp", groupnames=c("country", "age_years", "qcaeparent_tot"))

pain_years_country_qcae_summary %>%
  ggplot(aes(x=qcaeparent_tot, y=empathy_pain_resp, group=as.factor(age_years), color=as.factor(age_years))) + 
  geom_point()+
  geom_smooth(method = lm) +
  scale_color_brewer(palette="Paired")+theme_minimal() +
  #xlab("Age (years)") + ylab("Cognitive Empathy Score") +
  labs(colour = "Country") +
  facet_wrap(~country) +
  theme(plot.title = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.x = element_text(color = "grey20", size = 16, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 16, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 20, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 20, angle = 90, hjust = .5, vjust = .5, face = "plain"))

sorry_years_summary <- data_summary(empathy_data, varname="empathy_sorry_resp", groupnames=c("country", "age_years", "gender"))


# Summarize and visualize best-fitting models

plot_model(pain_years_country_qcae, type='int') 
tab_model(pain_years_country_qcae, show.std = 'std')
plot_model(pain_months_country_qcae, type='int') 
tab_model(pain_months_country_qcae, show.std = 'std')

plot_model(sorry_years_country_qcae, type='int') 
tab_model(sorry_years_country_qcae, show.std = 'std')
plot_model(sorry_months_country_qcae, type='int') 
tab_model(sorry_months_country_qcae, show.std = 'std')
```

#### Parent-Child Empathy Transmission
# Hypothesis: Self-reported parent empathy will mediate the country effect on empathy score
Dependent Variables: Behaviors (cognitive empathy and empathic concern scores)
Independent Variables: Hofstede collectivism score, parent total QCAE score
Analyses: For each country-behavior combination, run the following
1. lm(behavior ~ individualism)
* If b1 is significant, then there is a relationship between the parent’s self-reported empathy and the child’s behavior, which the country’s cultural environment may mediate
2. lm(QCAEparent_tot ~ individualism)
* If b2 is not significant, then the country’s QCAE is just a third variable which may or may not be associated with the behavior
3. lm(behavior ~ individualism + QCAEparent_tot)
* If b3 (QCAE) is significant and b4 (individualism) is not significant or smaller than before, then there is a mediation effect of culture on the relationship between parent empathy and child behavior.
4. mediate(lm(QCAEparent_tot ~ individualism, empathy_data), lm(behavior ~ individualism + QCAEparent_tot, empathy_data), treat=individualism, mediator= QCAEparent_tpt, boot=TRUE, sims=100)
* If the ACME score is significant, then the mediation effect is significant
5. lavaan(lm(behavior ~ individualism, empathy_data), lm(behavior ~ individualism + QCAEparent_tot, empathy_data), treat=individualism, mediator= QCAEparent_tpt, boot=TRUE, sims=100)
* Look for significant a, b, and ab effects

Interpretation: If there is a significant mediation effect of cultural individualism on the relationship between parent empathy and child behavior, then there are cross-cultural differences in transmission of empathy from parent to child

``` {r, echo = TRUE}
#Mediation by hand
# Note: for mediation to work, must remove NA values of outcome variable
pain_individualism <- lm(empathy_pain_resp_c ~ individualism, data = filter(empathy_data, !is.na(empathy_pain_resp_c)))
summary(pain_individualism) # b1 is significant, so there is a relationship between the parent’s self-reported empathy and the child’s behavior
sorry_individualism <- lm(empathy_sorry_resp_c ~ individualism, data = filter(empathy_data, !is.na(empathy_pain_resp_c)))
summary(sorry_individualism) # b1 is significant, so there is a relationship between the parent’s self-reported empathy and the child’s behavior

qcae_individualism <- lm(qcaeparent_tot ~ individualism, data = filter(empathy_data, !is.na(empathy_pain_resp_c)))
summary(qcae_individualism) # b2 is significant

pain_qcae_individualism <- lm(empathy_pain_resp_c ~ individualism + qcaeparent_tot, data = filter(empathy_data, !is.na(empathy_pain_resp_c)))
summary(pain_qcae_individualism) # b4 (individualism) is not insignificant or smaller than before, so mediation is unlikely
sorry_qcae_individualism <- lm(empathy_sorry_resp_c ~ individualism + qcaeparent_tot, data = filter(empathy_data, !is.na(empathy_pain_resp_c)))
summary(sorry_qcae_individualism) # b4 (individualism) is not insignificant or smaller than before, so mediation is unlikely


## Mediation using "mediation" package
model.M_pain <- lm(qcaeparent_tot ~ individualism, filter(empathy_data, !is.na(empathy_pain_resp_c)))
model.Y_pain <- lm(empathy_pain_resp ~ individualism + qcaeparent_tot, filter(empathy_data, !is.na(empathy_pain_resp_c)))
results_pain <- mediate(model.M_pain, model.Y_pain, treat='individualism', mediator='qcaeparent_tot',
                   boot=TRUE, sims=100)
summary(results_pain)
model.M_sorry <- lm(qcaeparent_tot ~ individualism, filter(empathy_data, !is.na(empathy_sorry_resp_c)))
model.Y_sorry <- lm(empathy_sorry_resp ~ individualism + qcaeparent_tot, filter(empathy_data, !is.na(empathy_sorry_resp_c)))
results_sorry <- mediate(model.M_sorry, model.Y_sorry, treat='individualism', mediator='qcaeparent_tot',
                   boot=TRUE, sims=100)
summary(results_sorry)



## Mediation using "lavaan" package

pain_data <- filter(data.frame(X = empathy_data$individualism, Y = empathy_data$empathy_pain_resp_c, M = empathy_data$qcaeparent_tot), !is.na(Y))

pain_fit <- lavaan::sem(pain_model, data = pain_data)
parameterestimates(pain_fit,  standardized=TRUE) %>% 
  knitr::kable()
sorry_data <- filter(data.frame(X = empathy_data$individualism, Y = empathy_data$empathy_sorry_resp_c, M = empathy_data$qcaeparent_tot), !is.na(Y))
sorry_model <- ' # direct effect
             Y ~ c*X
           # mediator
             M ~ a*X
             Y ~ b*M
           # indirect effect (a*b)
             ab := a*b
           # total effect
             total := c + (a*b)
         '
sorry_fit <- sem(sorry_model, data = sorry_data)
parameterestimates(sorry_fit,  standardized=TRUE) %>% 
  knitr::kable()


pain_model <- ' # direct effect
             Y ~ c*X
           # mediator
             M ~ a*X
             Y ~ b*M
           # indirect effect (a*b)
             ab := a*b
           # total effect
             total := c + (a*b)
         '



# Attempting Elizabeth's method
mediation_data <- filter(empathy_data, !is.na(empathy_pain_resp_c))

med_WB6 = "individualism  ~ a1*qcaeparent_tot
empathy_pain_resp_c ~ b1*individualism + c*qcaeparent_tot

indirect1 := a1*b1
direct := c
total := c + (a1*b1) 

"

fit_med_WB6 = lavaan:sem(med_WB6, se = "boot", bootstrap = 1000, data = mediation_data, 
                  likelihood = "wishart")
summary(fit_med_WB6, standardized = T, rsq = T)
parameterEstimates(fit_med_WB6)

# Attempting Elizabeth's method
mediation_data <- filter(empathy_data, !is.na(empathy_pain_resp_c))

med_WB6 = "individualism  ~ a1*qcaeparent_tot
empathy_pain_resp_c ~ b1*individualism + c*qcaeparent_tot

indirect1 := a1*b1
direct := c
total := c + (a1*b1) 

"
##EH notes - this is flipped, you want,
empathy_eh = "qcaeparent_tot  ~ a1*individualism
empathy_pain_resp_c ~ b1*qcaeparent_tot + c*qindividualism

indirect1 := a1*b1
direct := c
total := c + (a1*b1) 

"


fit_med_WB6 = lavaan::sem(med_WB6, se = "boot", bootstrap = 1000, data = mediation_data, 
                  likelihood = "wishart")
summary(fit_med_WB6, standardized = T, rsq = T)
parameterEstimates(fit_med_WB6)


```




