---
title: "Calculate Age"
author: "Violet Kozloff"
date: "June 8, 2020"
output: html_document
---

Set up workspace
```{r}
install.packages("here", "tidyverse", "lubridate")
require("here")
require("tidyverse")
require("lubridate")

rm(list = ls())
```
Read in data files
```{r}
canada_data <- as_tibble(read.csv(here("canada.csv")))
china_data_1 <- as_tibble(read.csv(here("china full qualtrics 1.csv")))
china_data_2 <- as_tibble(read.csv(here("china full qualtrics 2.csv")))
china_data_3 <- as_tibble(read.csv(here("china partial qualtrics 2.csv")))
jordan_data <- as_tibble(read.csv(here("jordan qualtrics.csv")))
south_africa_data_1 <- as_tibble(read.csv(here("south africa qualtrics_sheet1.csv")))
south_africa_data_2 <- as_tibble(read.csv(here("south africa qualtrics_sheet2.csv")))
turkey_data <- as_tibble(read.csv(here("turkish_mergeddata_allages.csv")))
usa_data <- as_tibble(read.csv(here("US Full Qualtrics.csv")))
```

Clean Canada data
```{r}
clean_canada_data <- canada_data %>% mutate (unique_id = paste0("canada_", ï..Subject))
clean_canada_data <- clean_canada_data %>% rename (qualtrics_age = age)
clean_canada_data <- clean_canada_data %>% mutate(collection_date = paste0(ifelse(todaysdate_mm<10, paste0("0", todaysdate_mm), todaysdate_mm), 
                                                                           "/", ifelse(todaysdate_dd<10, paste0("0", todaysdate_dd), todaysdate_dd),
                                                                           "/", todaysdate_yyyy))

clean_canada_data <- clean_canada_data %>% mutate(birth_date = paste0(ifelse(childbday_dd<10, paste0("0", childbday_dd), childbday_dd), 
                                                                           "/", ifelse(childbda_mm<10, paste0("0", childbda_mm), childbda_mm),
                                                                           "/", childbday_yyyy))
clean_canada_data <- clean_canada_data %>% select(unique_id, collection_date, birth_date, qualtrics_age)
clean_canada_data <- clean_canada_data %>% filter(unique_id != "canada_NA")
```
Clean China data
```{r}
clean_china_data_1 <- china_data_1 %>% mutate (unique_id = paste0("china_", ï..))
clean_china_data_1 <- clean_china_data_1 %>% rename (collection_date = X)
clean_china_data_1 <- clean_china_data_1 %>% rename (birth_date = X.143)
clean_china_data_1 <- clean_china_data_1 %>% slice(3:n())
clean_china_data_1 <- clean_china_data_1 %>% select(unique_id, collection_date, birth_date)
clean_china_data_1$qualtrics_age <- NA
clean_china_data_2 <- china_data_2 %>% mutate (unique_id = paste0("china_", ï..))
clean_china_data_2 <- clean_china_data_2 %>% rename (collection_date = X)
clean_china_data_2 <- clean_china_data_2 %>% rename (birth_date = X.143)
clean_china_data_2 <- clean_china_data_2 %>% slice(3:n())
clean_china_data_2 <- clean_china_data_2 %>% select(unique_id, collection_date, birth_date)
clean_china_data_2$qualtrics_age <- NA
clean_china_data_3 <- china_data_3 %>% mutate (unique_id = paste0("china_", ï..subjects.number))
clean_china_data_3 <- clean_china_data_3 %>% rename (collection_date = date.of.testing)
clean_china_data_3 <- clean_china_data_3 %>% rename (birth_date = date.of.birth)
clean_china_data_3 <- clean_china_data_3 %>% rename (qualtrics_age = Age..months.)
clean_china_data_3 <- clean_china_data_3 %>% select(unique_id, collection_date, birth_date, qualtrics_age)
clean_china_data <- rbind(clean_china_data_1, clean_china_data_2)
clean_china_data <- rbind(clean_china_data, clean_china_data_3)
clean_china_data <- filter(clean_china_data, unique_id!="china_")
```
Clean Jordan data
```{r}
clean_jordan_data <- jordan_data %>% mutate (unique_id = paste0("jordan_", ï..Subject))
clean_jordan_data <- clean_jordan_data %>% rename (collection_date = date)
clean_jordan_data <- clean_jordan_data %>% rename (birth_date = childbday)
clean_jordan_data <- clean_jordan_data %>% rename (qualtrics_age = age)
clean_jordan_data <- clean_jordan_data %>% select(unique_id, collection_date, birth_date, qualtrics_age)
clean_jordan_data <- clean_jordan_data %>% filter(unique_id != "jordan_NA")
```
Clean South Africa data
```{r}
clean_south_africa_data_1 <- south_africa_data_1 %>% mutate (unique_id = paste0("south_africa_", ï..Subject))
clean_south_africa_data_1 <- clean_south_africa_data_1 %>% rename (qualtrics_age = age)
clean_south_africa_data_1 <- clean_south_africa_data_1 %>% mutate(collection_date = paste0(ifelse(todaysdate_mm<10, paste0("0", todaysdate_mm), todaysdate_mm), "/", ifelse(todaysdate_dd<10, paste0("0", todaysdate_dd), todaysdate_dd), "/", todaysdate_yyyy))
clean_south_africa_data_1 <- clean_south_africa_data_1 %>% mutate(birth_date = paste0(ifelse(childbday_dd<10, paste0("0", childbday_dd), childbday_dd), "/", ifelse(childbday_mm<10, paste0("0", childbday_mm), childbday_mm), "/", childbday_yyyy))
clean_south_africa_data_1 <- clean_south_africa_data_1 %>% select(unique_id, collection_date, birth_date, qualtrics_age)

clean_south_africa_data_2 <- south_africa_data_2 %>% mutate (unique_id = paste0("south_africa_", ï..Q30))
clean_south_africa_data_2 <- clean_south_africa_data_2 %>% rename (qualtrics_age = X)
clean_south_africa_data_2 <- clean_south_africa_data_2 %>% slice(3:n())
clean_south_africa_data_2 <- clean_south_africa_data_2 %>% mutate(collection_date = paste0(ifelse(as.numeric(as.character(Q34_1_TEXT))<10, paste0("0", Q34_1_TEXT), as.character(Q34_1_TEXT)), "/", ifelse(as.numeric(as.character(Q34_2_TEXT))<10, paste0("0", Q34_2_TEXT), as.character(Q34_2_TEXT)), "/",  as.character(Q34_3_TEXT)))
clean_south_africa_data_2 <- clean_south_africa_data_2 %>% mutate(birth_date = paste0(ifelse(as.numeric(as.character(Q5_1_TEXT))<10, paste0("0", Q5_1_TEXT), as.character(Q5_1_TEXT)), "/", ifelse(as.numeric(as.character(Q5_2_TEXT))<10, paste0("0", Q5_2_TEXT), as.character(Q5_2_TEXT)), "/",  as.character(Q5_3_TEXT)))
clean_south_africa_data_2 <- clean_south_africa_data_2 %>% select(unique_id, collection_date, birth_date, qualtrics_age)
clean_south_africa_data <- rbind(clean_south_africa_data_1, clean_south_africa_data_2)
```
Clean Turkey data
```{r}
clean_turkey_data <- turkey_data %>% mutate (unique_id = paste0("turkey_", ï..Subject))
clean_turkey_data <- clean_turkey_data %>% rename (collection_date = todaysdate)
clean_turkey_data <- clean_turkey_data %>% rename (birth_date = childbday)
clean_turkey_data <- clean_turkey_data %>% rename (qualtrics_age = age)
clean_turkey_data <- clean_turkey_data %>% select(unique_id, collection_date, birth_date, qualtrics_age)
clean_turkey_data <- clean_turkey_data %>% filter(unique_id != "turkey_NA")
```
Clean USA data
```{r}
clean_usa_data <- usa_data %>% mutate (unique_id = paste0("usa_", ï..Subject))
clean_usa_data <- clean_usa_data %>% mutate(collection_date = paste0(ifelse(as.numeric(as.character(todaysdate_mm))<10, paste0("0", todaysdate_mm), as.character(todaysdate_mm)), "/", ifelse(as.numeric(as.character(todaysdate_dd))<10, paste0("0", todaysdate_dd), as.character(todaysdate_dd)), "/",  as.character(todaysdate_yyyy)))
clean_usa_data <- clean_usa_data %>% mutate(birth_date = paste0(ifelse(as.numeric(as.character(childbday_mm))<10, paste0("0", childbday_mm), as.character(childbday_mm)), "/", ifelse(as.numeric(as.character(childbday_dd))<10, paste0("0", childbday_dd), as.character(childbday_dd)), "/",  as.character(childbday_yyyy)))
clean_usa_data <- clean_usa_data %>% rename (qualtrics_age = age)
clean_usa_data <- clean_usa_data %>% select(unique_id, collection_date, birth_date, qualtrics_age)
clean_usa_data <- clean_usa_data %>% filter(unique_id != "usa_NA")
usa_data <- as_tibble(read.csv(here("US Full Qualtrics.csv")))
```
Combine data
```{r}
combined_data <- rbind(clean_canada_data, clean_china_data)
combined_data <- rbind(combined_data, clean_jordan_data)
combined_data <- rbind(combined_data, clean_south_africa_data)
combined_data <- rbind(combined_data, clean_turkey_data)
combined_data <- rbind(combined_data, clean_usa_data)
combined_data <- (as_tibble(combined_data))
```
Calculate ages
```{r}
combined_data <- combined_data %>% mutate(birth_date =ifelse(str_detect(combined_data$birth_date, "NA") | str_length(birth_date)<8, NA, birth_date))
# Get rid of entries with impossible birth dates
combined_data <- combined_data %>% mutate(birth_date =ifelse(birth_date=="04/31/2004" | birth_date == "14/02/2001", NA, birth_date))
combined_data <- combined_data %>% mutate(birth_date = as.Date(mdy(birth_date)))
combined_data <- combined_data %>% mutate(collection_date =ifelse(str_detect(combined_data$collection_date, "NA") | str_length(collection_date)<8, NA, collection_date))
combined_data <- combined_data %>% mutate(collection_date = as.Date(mdy(collection_date)))
# Calculate age in months
combined_data <- combined_data %>% mutate (calculated_age = floor( as.numeric(difftime(collection_date, birth_date, units ="days"))/(365.25/12)))
combined_data <- combined_data %>% mutate (calculated_age = floor(ifelse(is.na(calculated_age) | calculated_age<=0, as.numeric(qualtrics_age), calculated_age)))
```
Update ages for 2.1% of participants whose ages were calculated incorrectly by 1 month, possibly due to rounding errors or inconsistencies between leap years. 
These participants' ages in months were recalculated in Google Sheets using the DATEDIF function, then checked against https://www.calculator.net/age-calculator.html, and are individually corrected below.
```{r}
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "canada_1042", 97))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "canada_1201", 61))

combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "jordan_47", 98))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "jordan_56", 126))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "jordan_101", 114))
 
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2015", 111))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2020", 101))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2089", 104))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2114", 93))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2140", 108))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2152", 76))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2181", 154))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2216", 145))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "south_africa_2237", 167))

combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_44", 54))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_55", 61))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_109", 58))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_150", 61))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_156", 56))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_175", 37))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_436", 80))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_439", 76))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_539", 99))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_553", 103))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "turkey_570", 136))

combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "usa_62", 105))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "usa_235", 87))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "usa_247", 89))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "usa_321", 152))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "usa_457", 92))
combined_data <- combined_data %>% mutate(calculated_age=replace(calculated_age, unique_id == "usa_462", 129))
```
Clean combined data
```{r}
#Remove any duplicate rows
output_data <- combined_data %>% select(-qualtrics_age, -collection_date, -birth_date) %>% distinct()

# Uncomment to see any remaining duplicate entries with slightly different values
# View(output_data %>% add_count(unique_id) %>% filter(n>1))
  
# Remove duplicate entry with missing values from age data
output_data <- output_data %>% filter(!(is.na(calculated_age) & (unique_id == "china_3092")))

# Remove duplicate entries with conflicting values from age data
output_data <- output_data %>% filter(!(unique_id == "south_africa_2108" | unique_id == "south_africa_2116" | unique_id == "south_africa_4885"))
```

Calculate age in years and write to CSV
```{r}
output_data <- output_data %>% rename (vk_age_months = calculated_age)
output_data <- output_data %>% 
  mutate (vk_age_years = floor (vk_age_months/12)) %>% 
  select (unique_id, vk_age_years, vk_age_months)

# Remove any whose age cannot be calculated
output_data <- output_data %>% filter(!is.na(vk_age_months))

write.csv(output_data, "templeton_age_calculations.csv")

```
